---
title: "chords_bias"
author: "Jonathan Rosenblatt"
date: "13/08/2014"
output: pdf_document
---


```{r, echo=FALSE}
library(chords)
library(ggplot2)
library(parallel)
library(dplyr)
```


```{r}
getError <- function (beta, theta, sample.length) {
    dk <- c(2,10)
    true.Nks <- rep(0,max(dk)); true.Nks[dk] <- 1000
    
    true.log.bks <- rep(-Inf, max(dk))
    true.log.bks[dk] <- log(beta)+theta*log(dk)
    
    rds.simulated.object <- makeRdsSample(
      N.k =true.Nks , 
      b.k = exp(true.log.bks),
      sample.length = sample.length)
    rds.simulated.object$estimates <- estimate.b.k(rds.simulated.object)
    theta- getTheta(rds.simulated.object)$theta
}
# Testing
beta <- 5e-6
theta <- 0.1
sample.length <- 800L
getError(beta, theta, sample.length )
```



## Shape of distribution of theta:
```{r, cache=TRUE}
nsims <- 200
error <- replicate(nsims, getError(beta, theta, sample.length ))
qplot(error, geom="histogram")+geom_vline(aes(intercept=0), col='red')
```
Distribution seems fairly symmetric. We can thus concentrate on the firts two moments of the errors to study the performance of the estimator.



Replicate to compute bias:
```{r}
getBias <- function(nsims, beta, theta, sample.length){
  error <- replicate(nsims, getError(beta, theta, sample.length ))
  c(error=mean(error), sd=sd(error))
}
```




## Bias as function of sample lenth:
```{r, cache=TRUE}
nsims <- 10
lengths <- seq(1e2, 2e3, length.out = 20) %>% floor
biases <- mapply(getBias, nsims, beta, theta, lengths)
std.err <- biases['sd',]
error <- biases['error',]
qplot(x= lengths, y= error)+
  geom_segment(aes(x = lengths, y=error-2*std.err/nsims, 
                   xend=lengths, yend=error+2*std.err/nsims))+
  geom_hline(aes(itercept=0),lty=2)+ ylab('Bias')
```


## Bias as a function of theta:
```{r, cache=TRUE}
sample.length <- 800
thetas <- seq(-1,1, length.out = 11)
biases.2 <- mapply(getBias, nsims, beta, thetas, sample.length)

std.err <- biases.2['sd',]
error <- biases.2['error',]
qplot(x= thetas, y= error)+
  geom_segment(aes(x = thetas, y=error-std.err/nsims, 
                   xend=thetas, yend=error+std.err/nsims))+
    geom_hline(aes(itercept=0),lty=2)+ ylab('Bias')
```

## As a function of theta and sample length:
```{r, cache=TRUE}
nsims <- 50
lengths <- seq(1e2, 2e3, length.out = 7) %>% floor
thetas <- seq(-1,1, length.out = 7)

design <- expand.grid(nsims=nsims, beta=beta, theta=thetas, sample.length=lengths)
getBias.wrap <- function(x) do.call(getBias,as.list(x))

# Serial version
## biases.3 <- apply(design, 1, getBias.wrap)

# Paralle version:
cl <- makeCluster(detectCores())
clusterEvalQ(cl, {
  library(chords)
})
clusterExport(cl, c('getError','getBias'))
biases.3 <- parApply(cl, design, 1, getBias.wrap)
stopCluster(cl)

frame.3 <- data.frame(design, t(biases.3))
frame.3 <- mutate(frame.3, theta=round(theta,2))
qplot(data = frame.3,x=sample.length, y=error)+
  facet_grid(facets=.~theta, labeller = label_both)+
  geom_segment(aes(x=sample.length, y=error+2*sd/nsims, 
                   xend=sample.length, yend=error-2*sd/nsims))+
  scale_x_sqrt()+
  geom_hline(aes(intercept=0), lty=2)+ ylab('Bias')
```

Estimator is biased but clearly consistent. Convergence seems to be at $\sqrt{n}$. 


## As a function of theta, beta and sample length:
```{r, cache=TRUE}
nsims <- 20
lengths <- seq(1e2, 2e3, length.out = 7) %>% floor
thetas <- round(seq(-1,1, length.out = 5),2)
betas <- round(seq(1/6,1,length.out = 3),2)

design.2 <- expand.grid(nsims=nsims, beta=betas, theta=thetas, sample.length=lengths)
getBias.wrap <- function(x) do.call(getBias,as.list(x))

# Paralle version:
cl <- makeCluster(detectCores())
clusterEvalQ(cl, library(chords))
clusterExport(cl, c('getError','getBias'))
## Careful: long run:
biases.4 <- parApply(cl, design.2, 1, getBias.wrap)
stopCluster(cl)

frame.4 <- data.frame(design.2, t(biases.4))
frame.4 <- mutate(frame.4, theta=round(theta,2))
qplot(data = frame.4, x=sample.length, y=error)+
  facet_grid(facets=beta~theta, labeller = label_both)+
  geom_segment(aes(x=sample.length, y=error+2*sd/nsims, 
                   xend=sample.length, yend=error-2*sd/nsims))+
  scale_x_sqrt()+
  geom_hline(aes(intercept=0), lty=2)+ ylab('Bias')
```

Seems beta has no effect of the magnitude of the bias.